---
title: "Activity: Bird Species"
author: "Sritej Nadella"
date: "DUE DATE "
output: html_notebook
---


# Set Up

```{r}
# Load `tidyverse` and `dcData` packages
library(tidyverse)
library(mosaic)
library(esquisse)


# Load the `OrdwayBirds` data set into your RStudio environment from the `dcData` package
data("OrdwayBirds", package="dcData")
data("OrdwaySpeciesNames", package="dcData")



# Show a few commands used to inspect the data set a few different ways
summary('OrdwayBirds')


```

# Activity

When you're finished with this activity, you will have a graph that shows what time of year various species appear at the Katherine Ordway Natural History Study Area in Inver Grove Heights, MN.

### Step 0

Before we begin, the book gives us some instructions to select a few key variables and clean up the date formatting.  Use the commands given in the book, but be sure to study the functions used so you know what is happening.  In this case, we select `SpeciesName`, `Month`, and `Day`. Then a `mutate()` command converts each variable to character, and then converts the resulting "character string" to a number.

```{r}
# Get the data table & clean up dates (see Data Computing eBook)
OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month-as.numeric(as.character(Day)),
          Day = as.numeric(as.character(Day)))


# Inspect resulting data
head(OrdwayBirds)

```



### Step 1




```{r}
# unique species in original data (including mis-spellings)
OrdwayBirds %>%
  summarise( names=n_distinct( SpeciesName ))


# unique species in the clean list
OrdwaySpeciesNames %>%
  summarise( names=n_distinct( SpeciesNameCleaned ))


```


### Step 2
  



```{r}
# inspect OrdwaySpeciesNames
head(OrdwaySpeciesNames)

# inner join
Corrected <-
  OrdwayBirds %>%
  inner_join( OrdwaySpeciesNames ) %>%
  select( Species = SpeciesNameCleansed, Month, Day ) %>%
  na.omit()



# inspect the data after the join
head(Corrected)

```


### Step 3

We want to identify the top 5-6 bird species sighted at the Katherine Ordway Natural History Study Area.

```{r}
# count sightings by species in descending order
bird_sightings <-
  Corrected %>%
  group_by( Species ) %>%
  summarise( count = n() )

tall(bird_sightings)



# a hint in the text book recommends displaying the top 10 species to choose our threshold
Majors <-
  bird_sightings %>%
  arrange( desc(count) ) %>%
  slice( 1:10 )

head(Majors, 10)


# define a sightings threshold; produce a table with only birds belonging to major species
threshold <- 790
major_selected_birds <-
  Majors %>%
  filter( count >= threshold )

head(major_selected_birds)


```


### Step 4

We now want to produce a month-by-month count for each of the major species, and then construct a chart to tell the story of what time of year the various species appear.

```{r}
# prepare glyph-ready data table called `ByMonth`
ByMonth <-
  Corrected %>%
  group_by( Species, Month ) %>%
  summarise( count = n() ) %>%
  filter( Species %in% major_selected_birds$Species )
  


```

Now use the ggplot template below to generate the desired graph. (Remark you can also use esquisser or mplot to create a line chart instead of the barchart)



```{r}
# ggplot commands

ByMonth %>% 
  ggplot(aes(x = Month, y = count, color = Species)) + 
  geom_line()

ByMonth %>% ggplot(aes(x=Month,y=count ,fill=Species)) +
  geom_bar(stat='identity',position='stack', width=.9) 

```

According to the graph, we can answer the questions posed:

##### 1. Which species are present year-round?  
America Goldfinch and Black_capped chickadee


##### 2. Which species are migratory, that is, primarily present in one or two seasons?

Slate colored Junco


##### 3. What is the peak month for each major species? (bullet list)
.Bird month activity
.American Goldfinch = 10th month
.Black-capped chickadee = 10.9month
.Field Sparrow =  5th month
.Lincoln sparrow = 10th month
.Slate colored Junco = 10th month
.Tree Swallow= 3.73th month




##### 4. Which major species are seen in good numbers for at least 6 months of the year?
Ameican Goldfinch


```{r}

```

